---
title: "Analyzing Forest Fire Data"
author: "Namitha Deshpande"
date: "22/07/2020"
output: html_document
---

Data Source - https://archive.ics.uci.edu/ml/machine-learning-databases/forest-fires/

```{r}
#Loading packages
library(readr)
library(dplyr)
library(ggplot2)
library(magrittr)
library(purrr)
```

```{r}
#Loading our dataset
forest_data <- read.csv("forestfires.csv")
```
The "forestfires.csv" contains 517 rows and 13 columns in total. We will be performing exploratory analysis to better understand the data 

Creating a bar chart showing the number of forest fires occurring during each month to determine in which month are forest fires most common. 
```{r}
#obtaining month-wise forest fire data
month_fires <- forest_data %>%
  mutate(month = factor(month, levels = c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")))%>%
  group_by(month) %>%
  summarize(total_fires = n())

#bar-plot
ggplot(data = month_fires,
       aes(x = month, y = total_fires))+
  geom_bar(stat = "identity")+
  labs(title = "Forest Fires by Month")+
  theme(panel.background = element_rect(fill = "white"))
```

Creating a bar chart showing the number of forest fires occurring during each day of a week to determine which day of the week gets highest number of forest fires. 

```{r}
#Obtaining day-wise forest fires data
day_fires <- forest_data %>%
  mutate(day = factor(day, levels = c("sun", "mon", "tue", "wed", "thu", "fri", "sat")))%>%
  group_by(day) %>%
  summarize(total_fires = n())

#bar-plot
ggplot(data = day_fires) +
  aes(x = day, y = total_fires) +
  geom_bar(stat = "identity") +
  labs(title = "Forest Fires by Day of the Week") +
  theme(panel.background = element_rect(fill = "white"))
```

Creating boxplots for visualizing forest fires by month and day of the week

```{r}
## Write the function
create_boxplots <- function(x, y) {
  ggplot(data = forest_data) + 
    aes_string(x = x, y = y) +
    geom_boxplot() +
    theme(panel.background = element_rect(fill = "white"))
}

## Assign x and y variable names 
x_var_month <- names(forest_data)[3] ## month
x_var_day <- names(forest_data)[4] ## day
y_var <- names(forest_data)[5:12]

## use the map() function to apply the function to the variables of interest
month_box <- map2(x_var_month, y_var, create_boxplots) 
day_box <- map2(x_var_day, y_var, create_boxplots) 
```

Create scatter plots to see which variables may affect forest fire size: 

```{r}
## write the function 
create_scatterplots = function(x, y) {
  ggplot(data = forest_data) + 
    aes_string(x = x, y = y) +
    geom_point() +
    theme(panel.background = element_rect(fill = "white"))
}
## Assign x and y variable names 
x_var_scatter <- names(forest_data)[5:12]
y_var_scatter <- names(forest_data)[13]

## use the map() function to apply the function to the variables of interest
scatters <- map2(x_var_scatter, y_var_scatter, create_scatterplots)
```