---
title: "Association Rules"
author: "Namitha Deshpande"
date: "12/08/2020"
output: html_document
---

# **Association Rule Analysis on Titanic Data**

### **Data Description**

Get the data from http://www.rdatamining.com/data 

The sinking of the Titanic is one of the most infamous event in history, killing 1502 out of 2224 passengers and crew. This sensational tragedy shocked the international community and led to better safety regulations for ships. 

The Titanic dataset in the datasets package is a 4-dimensional table with summarized information on the fate of passengers on the Titanic according to social class, sex, age and survival.


```{r}
load("C:/Users/namit/Downloads/titanic.raw.rdata")
titanic_data <- titanic.raw
head(titanic_data)
```

### **Association Rule Mining**

An algorithm for association rule mining is APRIORI. It is a level-wise, breadth-first algorithm which counts transactions to find frequent itemsets and then derive association rules from them.

Below we demonstrate association rule mining with apriori(). With the function, the default settings are: 1) supp=0.1, which is the minimum support of rules; 2) conf=0.7, which is the minimum confidence of rules; and 3) maxlen=10, which is the maximum length of rules.

After association rule mining, rules are sorted by lift to make high-lift rules appear first.

```{r}
library(arules)
rules <- apriori(titanic.raw, control = list(verbose=F),
                 #min support & confidence
                 parameter = list(minlen=2, supp=0.005,conf=0.8),
                 appearance = list(rhs=c("Survived=No",
                                         "Survived=Yes"),
                                   default="lhs"))

quality(rules) <- round(quality(rules), digits=3)
rules.sorted <- sort(rules, by="lift")
inspect(rules.sorted)
```
### **Visualizing Association Rules**

Next we show some ways to visualize association rules, including scatter plot, balloon plot, graph and parallel coordinates plot.

```{r}
library(arulesViz)
plot(rules)
```

```{r}
plot(rules, method="graph")
```

```{r}
plot(rules, method="paracoord", control=list(reorder=TRUE))
```

### **Conclusion**

It seems that the number of children are less but proportion of there survived is more, compared to the adult. 

For female and adult, survival rate is highest based on the list value which represents proportion or probability of survived.

Even among the children-Female has more survival rate compare to the children-Male. 

