---
title: "Basic_COVID_Analysis"
author: "Namitha Deshpande"
date: "08/08/2020"
output: html_document
---
### **PACKAGE**
For this simple analysis, we will be using the **"Hmisc : Miscellaneous"** package. It Contains many functions useful for data analysis, high-level graphics, utility operations, functions for computing sample size and power, importing and annotating datasets, imputing missing values, advanced table making, variable clustering, character string manipulation, conversion of R objects to LaTeX and HTML code, and re-coding variables.

```{r}
library(Hmisc)
```

### **DATASET**

This analysis uses a Kaggle dataset - https://www.kaggle.com/sudalairajkumar/novel-corona-virus-2019-dataset

We will load and quickly scan the first few rows of the data.

```{r}
covid_data <- read.csv("C:/Users/namit/Desktop/COVID_R/datasets_494724_994000_COVID19_line_list_data.csv")
head(covid_data)
```

### **Cleaning the data**

```{r}
covid_data$death <- as.integer(covid_data$death != 0)
```

### **Simple Analysis**

Here, we will be using mainly age and gender attributes for our simple analysis

```{r}
#AGE
# claim: people who die are older
dead = subset(covid_data, death == 1)
alive = subset(covid_data, death == 0)
mean(dead$age, na.rm = TRUE)
mean(alive$age, na.rm = TRUE)

# is this statistically significant?
t.test(alive$age, dead$age, alternative="two.sided", conf.level = 0.99)
# normally, if p-value < 0.05, we reject null hypothesis
# here, p-value ~ 0, so we reject the null hypothesis and 
# conclude that this is statistically significant
```

This analysis showed that older people have greater chances of death compared to younger patients.

```{r}
# GENDER
# claim: gender has no effect
men = subset(covid_data, gender == "male")
women = subset(covid_data, gender == "female")
mean(men$death, na.rm = TRUE) #8.5%!
mean(women$death, na.rm = TRUE) #3.7%
# is this statistically significant?
t.test(men$death, women$death, alternative="two.sided", conf.level = 0.99)
# 99% confidence: men have from 0.8% to 8.8% higher chance
# of dying.
# p-value = 0.002 < 0.05, so this is statistically
# significant
```

In conclusion, we can say that men have higher chances of death compared to female patients
