---
title: "Diamonds_Data_EDA"
author: "Namitha Deshpande"
date: "16/08/2020"
output: html_document
---

# **Exploratory Data Analysis of `diamonds` Dataset**

In statistics, exploratory data analysis is an approach to analyzing data sets to summarize their main characteristics, often with visual methods. A statistical model can be used or not, but primarily EDA is for seeing what the data can tell us beyond the formal modeling or hypothesis testing task.

### **Dataset**

There are a total of 53,940 diamonds in the dataset with 10 features (carat, cut, color, clarity, depth, table, price, x,
y, and z)

We notice that the dataset is complete amd contains no missing values

```{r}
library(ggplot2)
data("diamonds")
summary(diamonds)
```

### **Summary Statistics**

Describe a dataset by summarizing its variables along 5 or more aspects such as central tendency, the distribution or the statistical dispersion. 

```{r}
#median and inter-quartile summary
fivenum(diamonds$price) 

#Boxplot hints towards possible outliers
boxplot(diamonds$price)
boxplot(diamonds$carat)
```

Check for outliers, check for their distance from the local extremes and obvious errors

### **Statistical Distribution**

Probability Distribution gives the possibility of occurrence of different events. It can be applied to both discrete and continuous variables.

The type of distribution of the dataset can be tested using visualization or hypothesis test.

```{r}
#Histogram to test the distribution
hist(diamonds$price)
```

We observe a positively or right skewed distribution for price variable with highest frequency between 0 and 5000 range. 

We will confirm that the price variable does not have normal distribution using `qqnorm()` function. If the distribution is normal, we will get a straight line, instead here we see a sigmoid curve.

```{r}
qqnorm(diamonds$price)
```

Same process can be applied to other variables as well

### **Hypothesis Testing for Distribution**

Under this, the data is compared against an ideal model. The data is tested to satify either null hypothesis or alternative hypothesis.

Shapiro-Wilk normality test is used here. Null Hypothesis states the data is normally distributed and alternative hypothesis rejects the null hypothesis. 

The deciding factor is p-value.If p-value >= 0.05 the null hypothesis is rejected and alternative hypothesis is rejected and vice versa.

```{r}
diam = as.data.frame(diamonds)
set.seed(50)
mysample = sample.int(n=53940, #total number of rows
                      size=5000, #selecting random sample
                      replace = F)
shapiro.test(diam[mysample,7]) #test for price
shapiro.test(diam[mysample,1]) #test for carat
```

Here the p-value is significant and well over 0.05, hence we reject null hypothesis.

### **Statistical Tests to find relation between variables**

We will first visualize relation between numeric data price and categorical data cut using box plot.

We observe that the median varies a bit, but the boxes lie in almost same range. Hence, we will use the Kruskal-Wallis test to check if there is significant difference between the groups based on p-value

The result says that each group either costs lower or higher than the other but does not say by how much. 

```{r}
#Relation b/w numeric and categorical data
boxplot(price ~ cut, data = diamonds)
kruskal.test(price ~ cut, data = diamonds)
```

Next we have relation between two numeric data. For this coefficient of correlation is applied. 

The result says that price and carat have strong positive relationship which means that as carat increases price increases and vice-versa.

```{r}
#Spearman method is used for skewed distribution
cor(diamonds$price, diamonds$carat,
    method = "spearman")
```

We can also view the relation by plotting a scatter plot as follows:

```{r}
plot(diamonds$price, diamonds$carat)
```

We can also generate a correlation matrix here with all the numeric variables involved.

```{r}
as.dist(cor(diamonds[,c(1,5:10)],
    method = "spearman"))
```

The correlation matrix can be plotted for easier understanding using the `heatmap()` function of R Base by tweaking few arguments as follows:

```{r}
library(GGally)

x = diamonds[,c(1,5:10)]
ggcorr(x, nbreaks = 20, mid = "white",
       high = "darkred", label = T,
       method = c("pairwise", "spearman"))
```

The last one is relation between categorical data which can be determined using the Chi-Squared Test. Here null hypothesis says no relation and alternative hypothesis states the opposite.

```{r}
mydata = table(diamonds[,c(2:4)]) #contingency table

chisq.test(diamonds$cut, diamonds$color)

chisq.test(diamonds$cut, diamonds$clarity)
```





