---
title: "Investigating COVID-19 Virus Trends"
author: "Namitha Deshpande"
date: "21/07/2020"
output: html_document
---

## Data Source - https://www.kaggle.com/lin0li/covid19testing

Our analysis in this project tries to provide an answer to this question: **Which countries have had the highest number of positive cases against the number of tests?**

My Goal with project is to test the basics to intermediate skills acquired while learning the R coursework. 

```{r}
library(tidyverse)
library(readr)

#loading the dataset
covid_df <- read_csv("covid19.csv")

#dimensions of the dataset
data_dimensions <- dim(covid_df)

#all the columns in our data
vector_cols <- colnames(covid_df)

#simple glance at our data
glimpse(covid_df)
```
The dataset contains 10903 rows and 14 columns in total. This database provides information on the numbers (per day and cumulatively) of COVID-19 positive cases, deaths, tests performed and hospitalizations for each country

```{r}
#removing inconsistent data column
covid_df_all_states <- covid_df %>%
  filter(Province_State == "All States") %>%
  select(-Province_State)
```

We can remove Province_State column without loosing information as we have filtered the dataset to contain information of "All States" in the previous step.

```{r}
#extracting daily based data from the dataset
covid_df_all_states_daily <- covid_df_all_states %>%
  select(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)
```

We have excluded the cumulative COVID columns and extracted only daily COVID calculations column 

```{r}
#summarising data columns
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>% 
  group_by(Country_Region) %>% 
  summarise(tested = sum(daily_tested), 
            positive = sum(daily_positive),
            active = sum(active),
            hospitalized = sum(hospitalizedCurr)) %>% 
  arrange(desc(tested))

#filtering Top 10
covid_top_10 <- head(covid_df_all_states_daily_sum, 10)

covid_top_10
```

```{r}
#obtaining vectors
countries <- covid_top_10$Country_Region
tested_cases <- covid_top_10$tested
positive_cases <- covid_top_10$positive
active_cases <- covid_top_10$active
hospitalized_cases <- covid_top_10$hospitalized
```

```{r}
#naming vectors
names(positive_cases) <- countries
names(tested_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries
```

```{r}
#Top 3 positive against tested
positive_cases
sum(positive_cases)
mean(positive_cases)
positive_cases/sum(positive_cases)

positive_cases/tested_cases

positive_tested_top_3 <- c("United Kingdom" = 0.11, "United States" = 0.10, "Turkey" = 0.08)
```

```{r}
# Creating vectors
united_kingdom <- c(0.11, 1473672, 166909, 0, 0)
united_states <- c(0.10, 17282363, 1877179, 0, 0)
turkey <- c(0.08, 2031192, 163941, 2980960, 0)

# Creating the matrix covid_mat
covid_mat <- rbind(united_kingdom, united_states, turkey)

# Naming columns
colnames(covid_mat) <- c("Ratio", "tested", "positive", "active", "hospitalized")

covid_mat
```

```{r}
#final analysis

question <- "Which countries have had the highest number of positive cases against the number of tests?"

answer <- c("Positive tested cases" = positive_tested_top_3)

datasets <- list(
  original = covid_df,
  allstates = covid_df_all_states,
  daily = covid_df_all_states_daily,
  top_10 = covid_top_10
)

matrices <- list(covid_mat)

vectors <- list(vector_cols, countries)

data_structure_list <- list("dataframe" = datasets, "matrix" = matrices, "vector" = vectors)

covid_analysis_list <- list(question, answer, data_structure_list)

covid_analysis_list[[2]]
```

